Viscous Transport in Eroding Porous Media

### 2019-07-07 docs/transport/erosion_transport.tex
add new figs in transport paper
small changes to paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Continuing with our previous work~\cite{qua-moo2018}, we use a
double-layer potential formulation for the velocity $\uu$.  The
double-layer potential is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) = 
    \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  In Section~\ref{sec:DLPcomplex}, we
show how the Stokes double-layer potential can also be expressed as the
sum of a Laplacian double-layer potential 
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \eta(\yy) ds_\yy
\end{align}
and its gradient.  Since the Laplacian double-layer potential can be
written as a Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$,
the Stokes double-layer potential can be written in terms of Cauchy
integrals and its first derivative.  We also require the derivative of
the Stokes double-layer potential which requires the second derivative
of Cauchy integrals.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Stokes double-layer potential~\eqref{eqn:stokesDLP} can be
decomposed as
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
All these integrals but the first are a gradient of a Laplace layer
potential.  In addition, each component of the first integral can be
written as a Cauchy integral with a complex-valued density
function~\cite{bar-wu-vee2015}.  Therefore, the Stokes double-layer
potential can be written as five Cauchy integrals and their derivatives
with the density functions $\yy \cdot \eeta$, $\eta_1$, and $\eta_2$.
The individual velocity components are
\begin{align}
  \label{eqn:cauchy1}
  u_1(x) &= \Re (v[\tau_1](x)) + \Re (v'[y\cdot\eta](x)) 
           -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
  u_2(x) &= \Re (v[\tau_2](x)) - \Im (v'[y\cdot\eta](x)) 
       +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
\end{align}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \tau_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad \text{and} \quad
  \tau_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex representation of the outward unit normal
$\nn \in \RR^2$

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  To characterize the
transport, we first require tracer trajectories (streamlines) that are
governed by
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0.
\end{align}
Instead of allowing the transport and erosion to happen simultaneously,
we freeze the geometry and compute the streamlines in this fixed
geometry.  Therefore, the velocity $\uu$ has no temporal dependence when
we characterize the transport.
{\color{red}
To study the interaction of fluid and fixed bodies in the porous media,
some parameters can be used to describle the interation quantitively,
for example the porosity, the permeability, and the tortuosity. 
The relation of porosity and tortuosity has been studied in~\cite{matyka2008tortuosity}.
The definition of porosity is the ratio 
$$\phi=\frac{A_v}{A_T}$$
where $A_v$ is the area of void-space and $A_T$ is the area of the calculating area.}
{\color{red} The tortuosity represents the weighted average travel distance of a tracer in a porous media. 
The definition of tortuosity is 
$$T=\frac{1}{d}\frac{\displaystyle\int_{S}u_x(y) \lambda(y)dy}{\displaystyle\int_{S}u_x(y)dy}$$
where $S$ is an arbitrary cross section which is perpendicular to the x axis,
$\lambda(y)$ is the length of streamline at y on $S$, 
$u_x(y)$ is the component of velocity at y which is normal to $S$, and 
$d$ is the distance from $S$ to the end cross section.
It depends on not only the geometric structure but also the velocity of the fluid 
which is different from the porosity since porosity only depends on the geometric structure.}
{\color{red}Beisdes, the local tortuosity~\cite{matyka2008tortuosity} has been used to analysis 
the local behavior of steamline around a body. 
The definition of local tortuosity of the tracers at $y$ on $S$ is
$$\tau(y)=\frac{\lambda(y)}{d}.$$ 
To calculate tortuosity $T$ numerically, we use the simple quaduature rule
$$T=\frac{1}{d}\frac{\displaystyle\sum_{i=1}^{N} u_x(y_i) \lambda(y_i) \Delta y_i}{\displaystyle\sum_{i=1}^{N} u_x(y_i) \Delta y_i},$$ 
where $y_i$ is the point on $S$ and $\Delta y_i = y_{i+1}-y_{i}$.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying Quadrature}
\label{sec:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  Since the shear
stress must only be computed once per time step, and the vorticity is
only used as a post-processing step, we accept the $\mathcal{O}(N^2)$
algorithm.  However, the velocity must be computed at every GMRES
iteration, which would make the overall method impractical without a
fast algorithm.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{50 nearly touching bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\color{red}
The second nearly touching bodies case is 50 bodies in a Hagen-Poiseuille flow.
}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Continuing with our previous work~\cite{qua-moo2018}, we use a
double-layer potential formulation for the velocity $\uu$.  The
double-layer potential is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) = 
    \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  In Section~\ref{s:DLPcomplex}, we show
how the Stokes double-layer potential can also be expressed as the sum
of a Laplacian double-layer potential 
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \eta(\yy) ds_\yy
\end{align}
and its gradient.  Since the Laplacian double-layer potential can be
written as a Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$,
the Stokes double-layer potential can be written in terms of Cauchy
integrals and its first derivative.  We also require the derivative of
the Stokes double-layer potential which requires the second derivative
of Cauchy integrals.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{s:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Stokes double-layer potential~\eqref{eqn:stokesDLP} can be
decomposed as
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
All these integrals but the first are a gradient of a Laplace layer
potential.  In addition, each component of the first integral can be
written as a Cauchy integral with a complex-valued density
function~\cite{bar-wu-vee2015}.  Therefore, the Stokes double-layer
potential can be written as five Cauchy integrals and their derivatives
with the density functions $\yy \cdot \eeta$, $\eta_1$, and $\eta_2$.
The individual velocity components are
\begin{align}
  \label{eqn:cauchy1}
  u_1(x) &= \Re (v[\tau_1](x)) + \Re (v'[y\cdot\eta](x)) 
           -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
  u_2(x) &= \Re (v[\tau_2](x)) - \Im (v'[y\cdot\eta](x)) 
       +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
\end{align}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \tau_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad \text{and} \quad
  \tau_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex representation of the outward unit normal
$\nn \in \RR^2$

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  To characterize the
transport, we first require tracer trajectories (streamlines) that are
governed by
\begin{align}
  \label{eqn:trajectories}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0.
\end{align}
Instead of allowing the transport and erosion to happen simultaneously,
we freeze the geometry and compute the streamlines in this fixed
geometry.  Therefore, the velocity $\uu$ has no temporal dependence when
we characterize the transport.

Key variables that are used to characterize porous media flow is the
porosity (area of void space), permeability, and tortuosity.  The
tortuosity, which  quantifies the amount of turning made by a
streamline, is
\begin{align*}
  T = \frac{1}{d}\frac{\displaystyle\int_{S}u_x(y)\lambda(y)\,dy}
  {\displaystyle\int_{S}u_x(y)\,dy},
\end{align*}
where $S$ is an cross-section perpendicular to the $x$ axis with length
$d$, $\lambda(y)$ is the length of the streamline that is initizalized
at the $y$-coordinate along $S$, and $u_x(y)$ is the $x$ component of
the velocity at the $y$-coordinate along $S$.  There have been efforts
to relate the tortuosity to the porosity~\cite{matyka2008tortuosity},
but these results can not be universal since the porosity which depends
on only a single number, the tortuosity is affected by the distribution
of pores within the porous network.  In addition to the hydraulic
tortuosity, a local tortuosity can be used to analyze the local behavior
of a streamline near a body~\cite{matyka2008tortuosity}.  The local
tortuosity of a streamline initialized at $y$ along $S$ is
\begin{align*}
  \tau(y) = \frac{\lambda(y)}{d}.
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying Quadrature}
\label{s:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  Since the shear
stress must only be computed once per time step, and the vorticity is
only used as a post-processing step, we accept the $\mathcal{O}(N^2)$
algorithm.  However, the velocity must be computed at every GMRES
iteration, which would make the overall method impractical without a
fast algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Streamlines and Tortuosity}
\label{s:streamlines}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Having developed numerically accurate methods to compute the velocity
$\uu$ at points close to the bodies, we are able to compute the
trajectories of particles passing through an eroded body.  To find the
trajectory governed by equation~\eqref{eqn:trajectories}, we use the
fourth-order explicit Runge-Kutta method with a fixed time step size.
Since our method is both high-order in time and space, we are able to
simulate dynamics very close to the eroding bodies.  Resolving the flow
in these small velocity regions is essential to accurately describe
anomalous diffusion and mixing~\cite{leb-den-dav-bol-car-dec-bou2011}.

{\color{red}
To calculate tortuosity $T$ numerically, we use the simple quaduature
rule
$$T=\frac{1}{d}\frac{\displaystyle\sum_{i=1}^{N} u_x(y_i) \lambda(y_i) \Delta y_i}{\displaystyle\sum_{i=1}^{N} u_x(y_i) \Delta y_i},$$ 
where $y_i$ is the point on $S$ and $\Delta y_i = y_{i+1}-y_{i}$.}

@@

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/anomalous_diffusion20d}
\caption{\label{fig:Eroding20anomalous} The analysis of anomalous diffusion in
 20 nearly touching bodies with respect to two prosities. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{50 nearly touching bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\color{red}
The second nearly touching bodies case is 50 bodies in a Hagen-Poiseuille flow.
}

@@

%To study the interaction of fluid and fixed bodies in the porous media,
%some parameters can be used to describle the interation quantitively,
%for example the porosity, the permeability, and the tortuosity. 
%The relation of porosity and tortuosity has been studied in~\cite{matyka2008tortuosity}.
%The definition of porosity is the ratio 
%$$\phi=\frac{A_v}{A_T}$$
%where $A_v$ is the area of void-space and $A_T$ is the area of the calculating area.}
%{\color{red} The tortuosity represents the weighted average travel distance of a tracer in a porous media. 
%The definition of tortuosity is 
%$$T=\frac{1}{d}\frac{\displaystyle\int_{S}u_x(y) \lambda(y)dy}{\displaystyle\int_{S}u_x(y)dy}$$
%where $S$ is an arbitrary cross section which is perpendicular to the x axis,
%$\lambda(y)$ is the length of streamline at y on $S$, 
%$u_x(y)$ is the component of velocity at y which is normal to $S$, and 
%$d$ is the distance from $S$ to the end cross section.
%It depends on not only the geometric structure but also the velocity of the fluid 
%which is different from the porosity since porosity only depends on the geometric structure.}
%{\color{red}Beisdes, the local tortuosity~\cite{matyka2008tortuosity} has been used to analysis 
%the local behavior of steamline around a body. 
%The definition of local tortuosity of the tracers at $y$ on $S$ is
%$$\tau(y)=\frac{\lambda(y)}{d}.$$ 
